Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-05 latex/main.tex
Updates from Shing

\subsection{Major deletions in the ARG}\label{sec:supplement_dels}
% Referenced by the main Results.
Basically, we want to say we did a lot of work to integrate signals of major
indels into the ARG, making it an even more powerful resource overall. Also,
indels are encoded by the Mutation-Annotated Tree format used by UShER
phylogenies (though PanMAT does this).

List of indels from Table S1 of Li et al. \cite{Li2023}
occurring at minimum frequency of 0.01 (out of 9,149,680 GISAID sequences passing filters) and 
their frequencies in the sc2ts ARG (maximum of max inheritors divided by 2,482,157 sampled sequences).

% See https://github.com/jeromekelleher/sc2ts-paper/issues/472
\begin{table}
\caption{Major indels.}
\label{tab:indels}
\begin{tabular}{rrcrrrr}
\toprule & \multicolumn{4}{c}{Li et al.} & \multicolumn{2}{c}{Sc2ts ARG} \\
\cmidrule(lr){1-5} \cmidrule(lr){6-7}
Start & Length & Region & Count & Frequency & Count & Frequency \\
\midrule
28362 & 9 & N & 3939164 & 0.4305 & 871774 & 0.3512 \\
28271 & 1 & Intergenic & 3853692 & 0.4212 & 1118363 & 0.4506 \\
21765 & 6 & S & 3179887 & 0.3475 & 339681 & 0.1368 \\
11288 & 9 & ORF1ab-nsp6 & 3040576 & 0.3323 & 290408 & 0.1170 \\
28248 & 6 & ORF8 & 3026186 & 0.3307 & 1114669 & 0.4491 \\
22029 & 6 & S & 3010665 & 0.3290 & 1112410 & 0.4482 \\
11283 & 9 & ORF1ab-nsp6 & 2011934 & 0.2199 & 40 & 0.0000 \\
6513 & 3 & ORF1ab-nsp3 & 1984683 & 0.2169 & 340765 & 0.1373 \\
21633 & 9 & S & 1941962 & 0.2122 & 532072 & 0.2144 \\
21987 & 9 & S & 1918094 & 0.2096 & N/A & N/A \\
22194 & 3 & S & 1833050 & 0.2003 & 312706 & 0.1260 \\
21991 & 3 & S & 912944 & 0.0998 & 288831 & 0.1164 \\
686 & 9 & ORF1ab-nsp1 & 115450 & 0.0126 & N/A & N/A \\
\bottomrule
\end{tabular}
\end{table}

>>>>

\subsection{Major deletions in the ARG}\label{sec:supplement_dels}
% Referenced by the main Results.
Li et al. \cite{Li2023} performed a large-scale analysis of indels 
using a global GISAID dataset (sampled up to August, 2022). 
We examined the major indels reported in Table S1 of that study 
which have a minimum frequency of 0.01 
(sequence count divided by 9,149,680 filtered sequences).
Most of these indels (all deletions) are encoded in the ARG (\ref{tab:indels}) 
and have high frequencies in the ARG
(sequence count divided by the total number of samples), 
except a 9-base deletion at 11283, a 9-based deletion at 21987, and a 9-base deletion at 686.
During site remapping, the 9-base deletion at 11283 was split up into 
a 5-base deletion at 11283 (count of 339867) and 
a 4-base deletion at 11288 (count of 871782).
The 686 deletion occurs below our threshold for inclusion into the ARG.

% See https://github.com/jeromekelleher/sc2ts-paper/issues/472
\begin{table}
\caption{Major deletions in Li et al. (2023).}
\label{tab:indels}
\begin{tabular}{rrcrrrr}
\toprule & \multicolumn{4}{c}{Li et al.} & \multicolumn{2}{c}{Sc2ts ARG} \\
\cmidrule(lr){1-5} \cmidrule(lr){6-7}
Start & Length & Region & Count & Frequency & Count & Frequency \\
\midrule
28362 & 9 & N & 3939164 & 0.4305 & 871774 & 0.3512 \\
28271 & 1 & Intergenic & 3853692 & 0.4212 & 1118363 & 0.4506 \\
21765 & 6 & S & 3179887 & 0.3475 & 339681 & 0.1368 \\
11288 & 9 & ORF1ab-nsp6 & 3040576 & 0.3323 & 290408 & 0.1170 \\
28248 & 6 & ORF8 & 3026186 & 0.3307 & 1114669 & 0.4491 \\
22029 & 6 & S & 3010665 & 0.3290 & 1112410 & 0.4482 \\
11283 & 9 & ORF1ab-nsp6 & 2011934 & 0.2199 & 40 & 0.0000 \\
6513 & 3 & ORF1ab-nsp3 & 1984683 & 0.2169 & 340765 & 0.1373 \\
21633 & 9 & S & 1941962 & 0.2122 & 532072 & 0.2144 \\
21987 & 9 & S & 1918094 & 0.2096 & N/A & N/A \\
22194 & 3 & S & 1833050 & 0.2003 & 312706 & 0.1260 \\
21991 & 3 & S & 912944 & 0.0998 & 288831 & 0.1164 \\
686 & 9 & ORF1ab-nsp1 & 115450 & 0.0126 & N/A & N/A \\
\bottomrule
\end{tabular}
\end{table}
